---
output:
  pdf_document: default
  html_document: default
  word_document: default
---
## Install necessary packages
```{r}
suppressMessages(library(devtools))
suppressMessages(library(ggplot2))
suppressMessages(library(ggthemes))
suppressMessages(library(dplyr))
suppressMessages(library(memisc))
suppressMessages(library(gridExtra))
suppressMessages(library(RColorBrewer))
suppressMessages(library(magrittr))
suppressMessages(library(xtable))
suppressMessages(library(knitr))
suppressMessages(library(DT))
suppressMessages(library(scales))
suppressMessages(library(plotrix))
suppressMessages(library(corrplot))
```



```{r echo=FALSE, message=FALSE, warning=FALSE}
loanData <- read.csv('prosperLoanData.csv')
```

```{r}
str(loanData)
```


## HOW LONG PEOPLE USUALLY OPT FOR LOAN?
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(sf)
attach(loanData)

loanData %>% 
    ggplot(aes(x = Term / 12)) +
        geom_histogram(binwidth = 1) +
        theme_hc() + 
        xlab('Term in Years') +
        ylab('How many?') + 
        scale_x_continuous(breaks = seq(1, 5, 2)) +
        ggtitle("Distribution Loan Terms") +
        theme(plot.title = element_text(face = 'bold.italic', 
                                        colour = "black", size=18))
```


## Plotting the trend of different customer types
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
loanData.two_status <- loanData %>%
    group_by(Term, LoanStatus) %>%
    summarise(n = n())
levels(loanData.two_status$LoanStatus) <-
    c(levels(loanData.two_status$LoanStatus), "Other")
loanData.two_status$LoanStatus[!(loanData.two_status$LoanStatus %in%
                                     c("Completed", "Defaulted"))] <- "Other"

loanData.two_status <- loanData.two_status %>%
    group_by(Term, LoanStatus) %>%
    summarise(p = n(), total = sum(n)) %>%
    mutate(freq = round(total / sum(total) * 100, 2))

ggplot(aes(x = Term / 12, y = freq, fill = LoanStatus), 
       data = loanData.two_status) +
    geom_bar(stat = 'identity', position="stack", color = 'black') + 
    scale_x_continuous(breaks = c(1, 3, 5)) +
    xlab('Loan term in years') + 
  
    ggtitle("LoanStatus: Completed vs Defaulted", 
            subtitle = "for each loan Term") +
    theme(plot.title = element_text(face = 'bold.italic', 
                                    colour = '#F35E3A', size=18),
          plot.subtitle = element_text(face = 'bold', 
                                       colour = '#17b4ba', size=11)) +
    ylab("% of Borrowers")
```


## Divide the customers into two groups-Good & Bad 
```{r echo=FALSE, message=FALSE, warning=FALSE}

loanData.gb <- loanData %>%
    group_by(Term, LoanStatus) %>%
    summarise(n = n()) %>%
    mutate(customer_type = ifelse((LoanStatus == 'Current' | 
                                   LoanStatus == 'Completed'| 
                                   LoanStatus == 'FinalPaymentInProgress'),
                                  'good', 'bad')) %>%
    filter(LoanStatus != 'Cancelled') %>%
    mutate(freq = n / sum(n) * 100) %>%
    ungroup() %>%
    group_by(Term, customer_type) %>%
    summarise(n = sum(freq))
```

## But before that lets see the distribution of LonaStatus variable
```{r echo=FALSE, message=FALSE, warning=FALSE}
loanData %>%
    group_by(LoanStatus) %>%
    summarise(n = n()) %>%
    ggplot(aes(x = LoanStatus, y = n)) +
    geom_bar(stat = 'identity', position="dodge") +
    ylab("Number of borrowers") +
    xlab("Different Loan Status") +
    coord_flip()
```

## The bigger Picture
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = Term / 12, y = n, fill = customer_type), data = loanData.gb) +
    geom_bar(stat = 'identity', position="dodge") + 
    xlab('Loan term in years') +
    ylab('Percentage') +
    scale_x_continuous(breaks = c(1, 3, 5)) +
    geom_text(aes(label = sprintf("%2.1f%%", round(n, 2)), vjust = -.3), 
              color="black") +
    facet_wrap(~customer_type) +
    scale_fill_hue() +
    ggtitle("Good vs Bad", 
            subtitle = "Number of good borrowes vs bad borrowers 
in each loan Term") +
    theme(plot.title = element_text(face = 'bold.italic', 
                                    colour = '#333338', 
                                    size=18),
          plot.subtitle = element_text(face = 'bold', 
                                       colour = '#6aa5e7', 
                                       size=11))
```


## Distribution for some continuous/categorical features

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(gridExtra)
library(ggplot2)
library(dplyr)

EmploymentStatusDurationHist <- ggplot(aes(x = EmploymentStatusDuration), 
                                       data = loanData) +
    geom_histogram(binwidth = 10, color = 'white') +
    scale_x_continuous(breaks = seq(0, 400, 100), limits = c(0, 400)) +
      xlab("Borrower Experience")


StatedMonthlyIncomeHist <- ggplot(aes(x = StatedMonthlyIncome), 
                                  data = loanData) +
    geom_histogram(binwidth = 1000, color = 'white') +
    scale_x_continuous(limits = c(0, 20000)) 

MonthlyLoanPaymentHist <- ggplot(aes(x = MonthlyLoanPayment), 
                                 data = loanData) +
    geom_histogram(binwidth = 50, color = 'white') +
    scale_x_continuous(breaks = seq(0, 1000, 300), limits = c(0, 1000)) 

LoanOriginalAmountHist <- ggplot(aes(x = LoanOriginalAmount), 
                                 data = loanData) +
    geom_histogram(binwidth = 1000, color = 'white') +
    scale_x_continuous(breaks = seq(0, 25000, 7000), limits = c(0, 25000)) 


BorrowerAPRHist <- ggplot(aes(x = BorrowerAPR), 
                          data = loanData) +
    geom_histogram(binwidth = 0.01, color = 'white') +
    scale_x_continuous(breaks = seq(0.07, 0.3, 0.05), limits = c(0.07, 0.3)) 

BorrowerRateHist <- ggplot(aes(x = BorrowerRate), 
                           data = loanData) +
    geom_histogram(binwidth = 0.01, color = 'white') +
    scale_x_continuous(breaks = seq(0.07, 0.3, 0.05), 
                       limits = c(0.07, 0.3)) 

DebtToIncomeRatioHist <- ggplot(aes(x = DebtToIncomeRatio), 
                                data = loanData) +
    geom_histogram(binwidth = 0.05, color = 'white') +
    scale_x_continuous(limits = c(0.0, 0.7), 
                       breaks = seq(0.0, 0.6, 0.1)) 

grid.arrange(EmploymentStatusDurationHist,
             StatedMonthlyIncomeHist,
             MonthlyLoanPaymentHist,
             LoanOriginalAmountHist,
             BorrowerAPRHist,
             BorrowerRateHist,
             DebtToIncomeRatioHist,
             ncol = 3, nrow = 3,
             top = "Distribution of feature variables\n")
    

```


## EstimatedEffectiveYield - A better measure for a successful Lender
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = EstimatedEffectiveYield), data = loanData) +
    geom_histogram(aes(y = ..density..), binwidth = 0.01, na.rm = T, 
                   color = 'darkblue', fill = 'lightblue') +
    scale_x_continuous(limits = c(-0.1, 0.5), 
                       breaks = seq(-0.1, 0.5, 0.05)) +
    geom_density(alpha=.2, fill="#FF6666", na.rm = T) + 
    geom_vline(aes(xintercept=mean(EstimatedEffectiveYield, na.rm=T)), 
               color="blue", linetype="dashed", size=1) +
    ggtitle("Distribution of EstimatedEffectiveYield", 
            subtitle = "with the mean axis") +
    theme(plot.title = element_text(face = 'bold.italic', 
                                    colour = '#FF6666', 
                                    size=18),
          plot.subtitle = element_text(face = 'bold', 
                                       colour = 'darkblue', 
                                       size=13))

```


## Does Lenders prefer borrowers with better Prosper Score ?
```{r echo=FALSE, message=FALSE, warning=FALSE}
loanData$ProsperScore <- factor(loanData$ProsperScore)
ggplot(aes(x = ProsperScore, y = EstimatedEffectiveYield, fill=ProsperScore), 
       data = subset(loanData, !is.na(loanData$ProsperScore) & 
                         !is.na(loanData$EstimatedEffectiveYield))) +
    geom_violin(trim = F, scale = "width") +
    stat_summary(fun.y=median, geom="point", size=2, color="black") +
    scale_fill_manual(values=colorRampPalette(c("pink", "lightgreen"))(11)) + 
    theme_minimal() +
    ylab('Effective yeild of Lenders') +
    ggtitle("Effective Yield for each Risk Factors",
            subtitle = "story of lenders preference") +
    theme(plot.title = element_text(face = 'bold.italic', 
                                    colour = '#f3b0c0', 
                                    size=22),
          plot.subtitle = element_text(face = 'bold', 
                                       colour = 'darkgreen', 
                                       size=14)) +
    guides(fill = F)
```

## we get an interesting trend
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=14}

borrower_apr <- ggplot(aes(x = ProsperScore, 
                           y = BorrowerAPR), 
       data = subset(loanData, !is.na(loanData$ProsperScore) & 
                         !is.na(loanData$BorrowerAPR))) +
    geom_boxplot() 

borrower_rate <- ggplot(aes(x = ProsperScore, 
                            y = BorrowerRate), 
       data = subset(loanData, !is.na(loanData$ProsperScore) & 
                         !is.na(loanData$BorrowerRate))) +
    geom_boxplot() 
grid.arrange(borrower_apr, borrower_rate, 
             nrow = 1, ncol = 2,
             top = "Interest Rate Distributions")
```

## Distribution of Listing Category
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=10}
loanData %>%
    group_by(ListingCategory..numeric.) %>%
    summarise(n = n()) %>%
    ggplot(aes(x = ListingCategory..numeric., y = n)) +
    geom_bar(stat = 'identity', position="dodge") +
    scale_x_continuous(breaks = seq(0, 20, 1)) +
    scale_y_continuous(breaks = seq(0, 60000, 5000)) +
    xlab("Listing Category") +
    ylab("Number of borrowers") +
    ggtitle("Distribution of ListingCategory")

```

## Who are others??
```{r echo=FALSE, message=FALSE, warning=FALSE}
other.borrowers <- loanData %>%
    filter(ListingCategory..numeric. == 7) %>%
    group_by(EmploymentStatus, Occupation) %>%
    summarise(n = n()) %>%
    filter(!is.na(Occupation)) %>%
    ungroup() %>%
    arrange(EmploymentStatus, desc(n)) %>%
    ungroup() %>%
    group_by(EmploymentStatus) %>%
    top_n(n = 5, wt = n) %>%
    ungroup()
```

## Does experienced people opt for loan lesser ?
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=11}
library(ggplot2)
library(gridExtra)
library(dplyr)
library(ggthemes)
borrower.experience <- loanData %>%
    filter(!is.na(EmploymentStatusDuration))

hist1 <- ggplot(aes(x = EmploymentStatusDuration / 12), 
                data = borrower.experience) +
         geom_histogram(binwidth = 1, 
                        color = 'red', 
                        fill = 'deeppink', 
                        alpha = 1/2) +
         theme_pander() +
         scale_x_continuous(breaks = 
                        seq(min(borrower.experience$EmploymentStatusDuration),
                        max(borrower.experience$EmploymentStatusDuration),
                                        2)) +
         coord_cartesian(xlim = c(0, 41)) +
         xlab('\nTotal experience in years') +
         ggtitle("Loans for Experienced people",
                 subtitle = "actual distribution") +
        theme(plot.title = element_text(face = 'bold.italic', 
                                        colour = 'deeppink', 
                                        size=20),
          plot.subtitle = element_text(face = 'bold', 
                                       colour = 'black', 
                                       size=14))

duration <- loanData %>%
    filter(!is.na(EmploymentStatusDuration)) %>%
    mutate(EmploymentStatusDuration = round(EmploymentStatusDuration / 12, 0)) %>%
    group_by(EmploymentStatusDuration) %>%
    summarise(n = n()) %>%
    mutate(freq = n / sum(n) * 100) %>%
    top_n(n = 5, wt = freq)


hist2<- 
    ggplot(aes(x = reorder(EmploymentStatusDuration, -freq), y = freq, fill = freq), 
           data = duration) +
        geom_bar(stat = 'identity', position="dodge",
             color = 'white') +
         theme_pander() +
    scale_fill_continuous_tableau(palette = "Blue") +
    guides(fill = F) +
    ggtitle("Loans for Experienced people",
                 subtitle = "top 5 experience levels") +
    theme(plot.title = element_text(face = 'bold.italic', 
                                        colour = 'darkslategrey', 
                                        size=20),
          plot.subtitle = element_text(face = 'bold', 
                                       colour = 'black', 
                                       size=14)) +
    xlab("Top 5 experience levels in years") +
    ylab("% of borrowers")
grid.arrange(hist1, hist2, ncol = 2, nrow = 1) 

```
    






## Other Correlations
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(corrplot)
d <- data.frame(Term=loanData$Term,
                BorrowerAPR=loanData$BorrowerAPR,
                BorrowerRate=loanData$BorrowerRate,
                LenderYield=loanData$LenderYield,
                EffectiveYield = loanData$EstimatedEffectiveYield,
                Loss = loanData$EstimatedLoss,
                Duration = loanData$EmploymentStatusDuration,
                income = loanData$StatedMonthlyIncome,
                payment = loanData$MonthlyLoanPayment)
d[is.na(d)] <- 0
d <- round(cor(d[sapply(d, is.numeric)]), 2)
corrplot(d, method = "circle")

```


## How people take loan for their Homes ?

Here we are going to explore the people for two category. 

    1. First those who are opting for loan to rennovation of home when they have a house.
    2. Second those who opt for home loans even though they don't have house.
```{r echo=FALSE, message=FALSE, warning=FALSE}
home.loans <- loanData %>%
    filter(ListingCategory..numeric. == 2, 
           ProsperRating..Alpha. != "", 
           IsBorrowerHomeowner != "") %>%
    group_by(IsBorrowerHomeowner, ProsperRating..Alpha.) %>%
    summarise(n = n()) %>%
    mutate(freq = round(n / sum(n) * 100, 2))
levels(home.loans$IsBorrowerHomeowner) <- c('No Home', 'Has Home')
home.loans$ProsperRating..Alpha. <- 
    factor(home.loans$ProsperRating..Alpha., ordered = T,
           levels = c('AA', 'A', 'B', 'C', 'D', 'E', 'HR'))
home.loans <- arrange(home.loans, ProsperRating..Alpha.)
```
**Now let's plot the data**

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = ProsperRating..Alpha., y = freq, 
           fill = IsBorrowerHomeowner), data = home.loans) +
    geom_bar(stat = 'identity', position="dodge",
             color = 'white') +
    scale_fill_hue() +
    xlab('Prosper Ratings for borrowers') +
    ylab('% of borrowers') +
    ggtitle("Home Improvement for All", 
            subtitle = "even if you dont have home") +
    theme(plot.title = element_text(face = 'bold.italic', 
                                    colour = 'red', 
                                    size=20),
          plot.subtitle = element_text(face = 'bold', 
                                       colour = 'dodgerblue3', 
                                       size=14))
```
